<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Stata Comparison: Linear Models}
-->

# Linear Models #

**margins** is intended as a port of (some of) the features of Stata's `margins` command. This vignette compares output from Stata's `margins` command for linear models against the output of **margins**.

```{r}
library("margins")
options(width = 100)
```

---

## OLS marginal effects ##

### Stata ###

```
. quietly reg mpg cyl hp wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)

------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.9416166   .5509165    -1.71   0.098    -2.070118    .1868846
          hp |  -.0180381   .0118763    -1.52   0.140    -.0423655    .0062893
          wt |  -3.166973    .740576    -4.28   0.000    -4.683975   -1.649972
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt

------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.9416166   .5509165    -1.71   0.098    -2.070118    .1868846
          hp |  -.0180381   .0118763    -1.52   0.140    -.0423655    .0062893
          wt |  -3.166973    .740576    -4.28   0.000    -4.683975   -1.649972
------------------------------------------------------------------------------
```

### R ###

```{r}
library("margins")
x <- lm(mpg ~ cyl + hp + wt, data = mtcars)
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```

---


## OLS marginal effects with interaction ##

### Stata ###

```
. quietly reg mpg cyl c.hp##c.wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.3652391   .5086204    -0.72   0.479    -1.408842    .6783638
          hp |  -.0252715   .0105097    -2.40   0.023    -.0468357   -.0037073
          wt |  -3.837584   .6730996    -5.70   0.000     -5.21867   -2.456498
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt

------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.3652391   .5086204    -0.72   0.479    -1.408842    .6783638
          hp |  -.0252715   .0105097    -2.40   0.023    -.0468357   -.0037073
          wt |  -3.837584   .6730996    -5.70   0.000     -5.21867   -2.456498
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- lm(mpg ~ cyl + hp * wt, data = mtcars)
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```

---

## OLS marginal effects with factor term ##

### Stata ###

```
. quietly reg mpg i.cyl hp wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : 6.cyl 8.cyl hp wt
at           : 4.cyl           =      .34375 (mean)
               6.cyl           =      .21875 (mean)
               8.cyl           =       .4375 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          6  |  -3.359024    1.40167    -2.40   0.024    -6.235014   -.4830353
          8  |  -3.185884   2.170476    -1.47   0.154    -7.639332    1.267564
             |
          hp |  -.0231198   .0119522    -1.93   0.064    -.0476437    .0014041
          wt |  -3.181404   .7196011    -4.42   0.000    -4.657904   -1.704905
------------------------------------------------------------------------------
Note: dy/dx for factor levels is the discrete change from the base level.


. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : 6.cyl 8.cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          6  |  -3.359024    1.40167    -2.40   0.024    -6.235014   -.4830353
          8  |  -3.185884   2.170476    -1.47   0.154    -7.639332    1.267564
             |
          hp |  -.0231198   .0119522    -1.93   0.064    -.0476437    .0014041
          wt |  -3.181404   .7196011    -4.42   0.000    -4.657904   -1.704905
------------------------------------------------------------------------------
Note: dy/dx for factor levels is the discrete change from the base level.
```

### R ###

```{r}
x <- lm(mpg ~ factor(cyl) + hp + wt, data = mtcars)
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```


---

## OLS marginal effects with squared term ##

### Stata ###

```
. quietly reg mpg cyl c.hp##c.hp wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.3696041   .6163571    -0.60   0.554    -1.634264    .8950561
          hp |  -.0429018   .0178353    -2.41   0.023    -.0794969   -.0063066
          wt |  -2.873553   .7301251    -3.94   0.001    -4.371646    -1.37546
------------------------------------------------------------------------------


. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.3696041   .6163571    -0.60   0.554    -1.634264    .8950561
          hp |  -.0429018   .0178353    -2.41   0.023    -.0794969   -.0063066
          wt |  -2.873553   .7301251    -3.94   0.001    -4.371646    -1.37546
------------------------------------------------------------------------------

```

### R ###

```{r}
x <- lm(mpg ~ cyl + hp + I(hp^2) + wt, data = mtcars)
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```

---

## OLS marginal effects with squared term (but no first-order term) ##

### Stata ###

```
. gen hp2 = hp^2
. quietly reg mpg cyl hp2 wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp2 wt
at           : cyl             =      6.1875 (mean)
               hp2             =    26071.19 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   -1.21919   .5030753    -2.42   0.022    -2.249693   -.1886869
         hp2 |   -.000028   .0000276    -1.01   0.320    -.0000846    .0000286
          wt |  -3.218637   .7570747    -4.25   0.000    -4.769435    -1.66784
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : cyl hp2 wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   -1.21919   .5030753    -2.42   0.022    -2.249693   -.1886869
         hp2 |   -.000028   .0000276    -1.01   0.320    -.0000846    .0000286
          wt |  -3.218637   .7570747    -4.25   0.000    -4.769435    -1.66784
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- lm(mpg ~ cyl + I(hp^2) + wt, data = mtcars)
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```

