<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Stata Comparison: Generalized Linear Models}
-->

# Generalized Linear Models #

**margins** is intended as a port of (some of) the features of Stata's `margins` command. This vignette compares output from Stata's `margins` command for generalized linear models against the output of **margins**.

```{r}
library("margins")
```

---

## GLM (Logit) predicted values ##

### Stata ###

```
. import delimited mtcars.csv
. quietly logit am cyl hp wt
* probability of success
. margins 

Predictive margins                                Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       _cons |     .40625   .0370796    10.96   0.000     .3335752    .4789248
------------------------------------------------------------------------------

* Linear prediction
. margins, predict(xb) 

Predictive margins                                Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       _cons |   -1.93545   1.144232    -1.69   0.091    -4.178104    .3072045
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp + wt, data = mtcars, family=binomial)
mean(predict(x, type = "response")) # Probability of success
mean(predict(x, type = "link")) # linear prediction
```

## GLM (Logit) predicted values at specified covariate levels ##


### Stata ###

```
* average predictive margins
. margins, at(cyl = (4 6 8))

Predictive margins                                Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
1._at        : cyl             =           4
2._at        : cyl             =           6
3._at        : cyl             =           8
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         _at |
          1  |   .3618736   .1065746     3.40   0.001     .1529913     .570756
          2  |   .4022889   .0363634    11.06   0.000      .331018    .4735599
          3  |    .439427   .0671516     6.54   0.000     .3078122    .5710418
------------------------------------------------------------------------------

* holding all other variables at means
. margins, at(cyl = (4 6 8)) atmeans 

Adjusted predictions                              Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
1._at        : cyl             =           4
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
2._at        : cyl             =           6
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
3._at        : cyl             =           8
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         _at |
          1  |   .0473325   .1307238     0.36   0.717    -.2088814    .3035464
          2  |   .1164102   .1254196     0.93   0.353    -.1294077    .3622281
          3  |   .2589031   .3718624     0.70   0.486    -.4699337    .9877399
------------------------------------------------------------------------------
```


### R ###

```{r}
x <- glm(am ~ cyl + hp + wt, data = mtcars, family=binomial)
# average predictive margins
sapply(c(4,6,8), function(z) {
    mean(predict(x, newdata = `[<-`(x$model, , "cyl", z), type = "response"))
})

# holding all other variables at their means
n <- expand.grid(cyl = c(4,6,8), hp = mean(mtcars$hp), wt = mean(mtcars$wt))
predict(x, newdata = n, type = "response")
```

---

## GLM (Logit) predicted values with interactions ##

### Stata ###

```
. quietly logit am i.cyl##c.hp wt
. margins

Predictive margins                                Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       _cons |   .4062558    .031352    12.96   0.000     .3448069    .4677046
------------------------------------------------------------------------------

. margins cyl

Predictive margins                                Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
------------------------------------------------------------------------------
             |            Delta-method
             |     Margin   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          4  |   .4475668   .1511923     2.96   0.003     .1512353    .7438982
          6  |   .4545709   .3008264     1.51   0.131    -.1350381     1.04418
          8  |      .0625   1.96e-08  3.2e+06   0.000        .0625       .0625
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ factor(cyl) * hp + wt, data = mtcars, family=binomial)
mean(predict(x, type = "response"))

sapply(c(4,6,8), function(z) {
    mean(predict(x, newdata = `[<-`(x$model, , "cyl", z), type = "response"))
})
```

---

## GLM (Logit) Effects on Probability Scale ##

### Stata ###

```
. quietly logit am cyl hp wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .0537505   .1132654     0.47   0.635    -.1682456    .2757465
          hp |   .0035928   .0029037     1.24   0.216    -.0020983    .0092838
          wt |  -1.008594   .6676631    -1.51   0.131     -2.31719    .3000017
------------------------------------------------------------------------------


. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .0214527   .0469746     0.46   0.648    -.0706157    .1135212
          hp |   .0014339   .0006182     2.32   0.020     .0002224    .0026455
          wt |  -.4025475   .1154098    -3.49   0.000    -.6287466   -.1763484
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp + wt, data = mtcars, family = binomial)
# MEM
margins(x, atmeans = TRUE, type = "response")
# AME
margins(x, type = "response")
```

---

## GLM (Logit) Effects on Log-Odds Scale ##


### Stata ###

```
. quietly logit am cyl hp wt
. margins, dydx(*) atmeans predict(xb)

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .4875978   1.071621     0.46   0.649    -1.612741    2.587936
          hp |   .0325917   .0188611     1.73   0.084    -.0043753    .0695587
          wt |   -9.14947   4.153326    -2.20   0.028    -17.28984   -1.009101
------------------------------------------------------------------------------

. margins, dydx(*) predict(xb)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .4875978   1.071621     0.46   0.649    -1.612741    2.587936
          hp |   .0325917   .0188611     1.73   0.084    -.0043753    .0695587
          wt |   -9.14947   4.153326    -2.20   0.028    -17.28984   -1.009101
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp + wt, data = mtcars, family = binomial)
# MEM
margins(x, atmeans = TRUE, type = "link")
# AME
margins(x, type = "link")
```


---

## GLM (Logit) Effects with factor variable on Log-Odds and probability scales ##


### Stata ###

```
. quietly logit am i.cyl hp wt
. margins, dydx(*) predict(xb)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
dy/dx w.r.t. : 6.cyl 8.cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          6  |   2.765754   3.156829     0.88   0.381    -3.421517    8.953025
          8  |  -8.388958   13.16745    -0.64   0.524     -34.1967    17.41878
             |
          hp |    .103209   .0960655     1.07   0.283    -.0850759    .2914939
          wt |  -10.67598   5.441998    -1.96   0.050     -21.3421   -.0098575
------------------------------------------------------------------------------
Note: dy/dx for factor levels is the discrete change from the base level.

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : 6.cyl 8.cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          6  |   .1197978   .1062873     1.13   0.260    -.0885214    .3281171
          8  |  -.3478575   .2067542    -1.68   0.092    -.7530883    .0573732
             |
          hp |   .0033268   .0029852     1.11   0.265    -.0025241    .0091777
          wt |  -.3441297   .1188604    -2.90   0.004    -.5770919   -.1111675
------------------------------------------------------------------------------
Note: dy/dx for factor levels is the discrete change from the base level.

. margins, dydx(*) continuous

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : 6.cyl 8.cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |
          6  |   .0891514   .0940577     0.95   0.343    -.0951984    .2735011
          8  |  -.2704099   .4252041    -0.64   0.525    -1.103795    .5629749
             |
          hp |   .0033268   .0029852     1.11   0.265    -.0025241    .0091777
          wt |  -.3441297   .1188604    -2.90   0.004    -.5770919   -.1111675
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ factor(cyl) + hp + wt, data = mtcars, family = binomial)
# Log-odds
margins(x, type = "link")
# Probability
margins(x, type = "response")
# Probability with discrete change
#margins(x, type = "response") # <- NEED AN OPTION TO CONTROL CONTINUOUS/DISCRETE CHANGE
```


---

## GLM (Logit) with interaction on probability scale ##

### Stata ###

```
. quietly logit am cyl c.hp##c.wt
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .0810915   .1838916     0.44   0.659    -.2793294    .4415125
          hp |   .0081009   .0086664     0.93   0.350    -.0088849    .0250867
          wt |  -1.925325   1.866456    -1.03   0.302    -5.583512    1.732861
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .0215633   .0492676     0.44   0.662    -.0749994    .1181261
          hp |   .0026673   .0023004     1.16   0.246    -.0018414     .007176
          wt |  -.5157922   .2685806    -1.92   0.055    -1.042201    .0106162
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp * wt, data = mtcars, family = binomial)
# MEM
margins(x, atmeans = TRUE, type = "response")
# AME
margins(x, type = "response")
```

---

## GLM (Logit) with interaction on Log-Odds scale ##

### Stata ###

```
. quietly logit am cyl c.hp##c.wt
. margins, dydx(*) atmeans predict(xb)

Conditional marginal effects                      Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
dy/dx w.r.t. : cyl hp wt
at           : cyl             =      6.1875 (mean)
               hp              =    146.6875 (mean)
               wt              =     3.21725 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .5156396   1.169458     0.44   0.659    -1.776456    2.807735
          hp |   .0515116    .035699     1.44   0.149    -.0184571    .1214804
          wt |  -12.24264   7.678428    -1.59   0.111    -27.29208    2.806807
------------------------------------------------------------------------------

. margins, dydx(*) predict(xb)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction (log odds), predict(xb)
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .5156396   1.169458     0.44   0.659    -1.776456    2.807735
          hp |   .0515116    .035699     1.44   0.149    -.0184571    .1214804
          wt |  -12.24264   7.678428    -1.59   0.111    -27.29208    2.806807
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp * wt, data = mtcars, family = binomial)
# MEM
margins(x, atmeans = TRUE, type = "link")
# AME
margins(x, type = "link")
```


## GLM (Probit) Effects on Log-Odds and probability scales ##

### Stata ###

```
. quietly probit am cyl c.hp##c.wt
. margins, dydx(*) predict(xb)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction, predict(xb)
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |   .2974758   .6629205     0.45   0.654    -1.001825    1.596776
          hp |   .0277713   .0193121     1.44   0.150    -.0100797    .0656223
          wt |  -6.626949   4.096208    -1.62   0.106    -14.65537    1.401471
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Pr(am), predict()
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |    .022611   .0498253     0.45   0.650    -.0750447    .1202667
          hp |   .0025769   .0022607     1.14   0.254     -.001854    .0070077
          wt |   -.508829   .2625404    -1.94   0.053    -1.023399    .0057408
------------------------------------------------------------------------------
```

### R ###

```{r}
x <- glm(am ~ cyl + hp * wt, data = mtcars, family = binomial(link="probit"))
# AME (log-odds)
margins(x, type = "link")
# AME (probability)
margins(x, type = "response")
```


## GLM (Poisson) Effects on Log-Odds and probability scales ##

### Stata ###

```
. quietly poisson carb cyl c.hp##c.wt
. margins, dydx(*) predict(xb)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Linear prediction, predict(xb)
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.0993854   .1478936    -0.67   0.502    -.3892516    .1904808
          hp |   .0066519   .0024217     2.75   0.006     .0019054    .0113984
          wt |   .1225051   .2035185     0.60   0.547    -.2763837     .521394
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =         32
Model VCE    : OIM
Expression   : Predicted number of events, predict()
dy/dx w.r.t. : cyl hp wt
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
         cyl |  -.2795214   .4169931    -0.67   0.503    -1.096813      .53777
          hp |   .0175935   .0067179     2.62   0.009     .0044267    .0307604
          wt |   .2075447   .4859868     0.43   0.669    -.7449719    1.160061
------------------------------------------------------------------------------
```


### R ###

```{r}
x <- glm(carb ~ cyl + hp * wt, data = mtcars, family = poisson)
# AME (linear/link)
margins(x, type = "link")
# AME (probability)
margins(x, type = "response")
```
