<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Stata Comparison: Panel Models}
-->

# Panel Regression Models #

**margins** is intended as a port of (some of) the features of Stata's `margins` command. This vignette compares output from Stata's `margins` command for panel models against the output of **margins**.


```{r}
library("margins")
```

---

## Linear Panel Model (Pooling) ##

### Stata ###

```
. import delimited Cigar.csv
. xtset state year
. quietly reg sales price pop cpi
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |   -.818001   .0577155   -14.17   0.000    -.9312209   -.7047811
         pop |  -.0003927   .0001585    -2.48   0.013    -.0007037   -.0000817
         cpi |   .7182534   .0661716    10.85   0.000     .5884452    .8480615
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |   -.818001   .0577155   -14.17   0.000    -.9312209   -.7047811
         pop |  -.0003927   .0001585    -2.48   0.013    -.0007037   -.0000817
         cpi |   .7182534   .0661716    10.85   0.000     .5884452    .8480615
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)
x <- plm(sales ~ price + pop + cpi, data = Cigar, model = "pooling")
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```


---

## Linear Fixed Effects Panel Model ##

### Stata ###

```
. xtset state year
. quietly xtreg sales price pop cpi, fe
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.6994539   .0302887   -23.09   0.000    -.7588187    -.640089
         pop |   .0001777   .0005044     0.35   0.725    -.0008108    .0011663
         cpi |   .5827791   .0346912    16.80   0.000     .5147857    .6507726
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.6994539   .0302887   -23.09   0.000    -.7588187    -.640089
         pop |   .0001777   .0005044     0.35   0.725    -.0008108    .0011663
         cpi |   .5827791   .0346912    16.80   0.000     .5147857    .6507726
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)
x <- plm(sales ~ price + pop + cpi, data = Cigar, model = "within", index = c("state","year"))
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
x <- lm(sales ~ price + pop + cpi + factor(state), data = Cigar)
# MEM
summary(margins(x, atmeans = TRUE))[1:3,]
# AME
summary(margins(x))[1:3,]
```

---

## Linear Random Effects Panel Model ##

### Stata ###

```
. xtset state year
. quietly xtreg sales price pop cpi
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |   -.700356   .0302552   -23.15   0.000    -.7596551   -.6410569
         pop |  -2.44e-06   .0004238    -0.01   0.995    -.0008331    .0008283
         cpi |   .5857053   .0345976    16.93   0.000     .5178952    .6535154
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |   -.700356   .0302552   -23.15   0.000    -.7596551   -.6410569
         pop |  -2.44e-06   .0004238    -0.01   0.995    -.0008331    .0008283
         cpi |   .5857053   .0345976    16.93   0.000     .5178952    .6535154
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)
x <- plm(sales ~ price + pop + cpi, data = Cigar, model = "random", index = c("state","year"))
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```



---

## Linear Panel Model (Pooling) with interaction ##

### Stata ###

```
. xtset state year
. quietly reg sales price c.pop##c.cpi
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.8221654   .0579823   -14.18   0.000    -.9359088    -.708422
         pop |  -.0004153   .0001613    -2.57   0.010    -.0007318   -.0000989
         cpi |   .7226397   .0664315    10.88   0.000     .5923217    .8529577
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : OLS
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      t    P>|t|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.8221654   .0579823   -14.18   0.000    -.9359088    -.708422
         pop |  -.0004153   .0001613    -2.57   0.010    -.0007318   -.0000989
         cpi |   .7226397   .0664315    10.88   0.000     .5923217    .8529577
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)
x <- plm(sales ~ price + pop * cpi, data = Cigar, model = "pooling")
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```


---

## Linear Fixed Effects Panel Model with interaction ##

### Stata ###

```
. xtset state year
. quietly xtreg sales price c.pop##c.cpi, fe
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.7016593   .0304284   -23.06   0.000    -.7612978   -.6420208
         pop |  -.0001637   .0006715    -0.24   0.807    -.0014799    .0011525
         cpi |   .5886036   .0355108    16.58   0.000     .5190037    .6582035
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.7016593   .0304284   -23.06   0.000    -.7612978   -.6420208
         pop |  -.0001637   .0006715    -0.24   0.807    -.0014799    .0011525
         cpi |   .5886036   .0355108    16.58   0.000     .5190037    .6582035
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)

# `plm` notation
x <- plm(sales ~ price + pop * cpi, data = Cigar, model = "within", index = c("state", "year"))
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)

# `lm` notation
x <- lm(sales ~ price + pop * cpi + factor(state), data = Cigar)
# MEM
summary(margins(x, atmeans = TRUE))[1:3,]
# AME
summary(margins(x))[1:3,]
```


---

## Linear Random Effects Panel Model with interaction ##

### Stata ###

```
. xtset state year
. quietly xtreg sales price c.pop##c.cpi
. margins, dydx(*) atmeans

Conditional marginal effects                      Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
at           : price           =    68.69993 (mean)
               pop             =    4537.113 (mean)
               cpi             =    73.59667 (mean)
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.7031592   .0303831   -23.14   0.000     -.762709   -.6436094
         pop |  -.0002812   .0005135    -0.55   0.584    -.0012876    .0007253
         cpi |   .5914665   .0350954    16.85   0.000     .5226809    .6602522
------------------------------------------------------------------------------

. margins, dydx(*)

Average marginal effects                          Number of obs   =       1380
Model VCE    : Conventional
Expression   : Linear prediction, predict()
dy/dx w.r.t. : price pop cpi
------------------------------------------------------------------------------
             |            Delta-method
             |      dy/dx   Std. Err.      z    P>|z|     [95% Conf. Interval]
-------------+----------------------------------------------------------------
       price |  -.7031592   .0303831   -23.14   0.000     -.762709   -.6436094
         pop |  -.0002812   .0005135    -0.55   0.584    -.0012876    .0007253
         cpi |   .5914665   .0350954    16.85   0.000     .5226809    .6602522
------------------------------------------------------------------------------
```

### R ###

```{r}
library("plm")
data(Cigar)
x <- plm(sales ~ price + pop * cpi, data = Cigar, model = "random", index = c("state","year"))
# MEM
margins(x, atmeans = TRUE)
# AME
margins(x)
```

